---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# eqc

It implements the ensemble quantile classifier (EQC) that extends the quantile-based classifier (Hennig and Viroli, 2016b).
An accompanying vignette illustrates a simulation study and the application of
the EQC compared to the other methods.

## Installation

To install this package, run

```{r,eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("CliffordLai/eqc")

# To include the vignette, one can use the following 
# but it takes much more time to install the package
devtools::install_github("CliffordLai/eqc",build_vignettes = TRUE)
```

## Example

This is a basic example which shows you how to train and tune the EQC.

```{r example}
library(eqc)
# Divide data into training set and test set randomly
data(wdbc)
set.seed(193)
trainIndex <- sample(c(rep(TRUE,2),rep(FALSE,1)),nrow(wdbc),replace=TRUE)
train <- as.matrix(wdbc[trainIndex,-1])
cl.train <- wdbc[trainIndex,1]
test <- as.matrix(wdbc[!trainIndex,-1])
cl.test <- wdbc[!trainIndex,1]
p <- ncol(train)

# Tuning parameters
thetaList <- rbind(rep(0.3,p),rep(0.4,p),rep(0.5,p))
# Fit the tuned EQC with a ridge logistic regression
lambda <- c(3,1,0.1,0.01,0.001,0.0005,0.0001)
fit <- eqcTrain(train,cl.train,
                 thetaList=thetaList,
                 method = "glmnet",
                 alpha = 0,lambda = lambda,
                 tuneControl = list(nfolds=5,fold.seed=117),
                 lower.limits=0, upper.limits=Inf)
fit$cvparameter #Selected tuning parameters
acc <- mean(predict(fit,newdata = test,type = "class")[[1]]==cl.test)
acc  #Classification accuracy

```


## Vignette

The vignette include a simulation study and an application. 
List vignettes in an HTML browser by,

```{r,eval=FALSE}
browseVignettes(package = "eqc")
```

