---
title: "Application of EQC on Text Data"
author: "Yuanhao"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: EQC_July22.bib
vignette: >
  %\VignetteIndexEntry{Application of EQC on Text Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In this vignette, 
we use a subset of the Reuters-21578 text categorization test collection [@Lewis1997] 
to demonstrate the usefulness of EQC. 

The following classifiers are compared:

* QC: Quantile-based Classifier [@Hennig2016];
* MC: Median-based classifier [@HALL2009];
* EQC/LSVM: EQC with coefficients fitted by the linear SVM;
* EQC/Ridge: EQC with coefficients fitted by the ridge logistic regression;
* EQC/Lasso: EQC with coefficients fitted by the Lasso logistic regression.

```{r, warning=FALSE,message=FALSE}
library(eqc)
library(tm)
library(NLP)
library(SnowballC)
```

## Data Preprocess 
The subset contains two topics, “acq” and “crude”, which can be found from the R package
tm [@tm]. The subset have 70 observations (documents),
where \eqn{n1 = 50} are of the topic “acq” and \eqn{n2 = 20} are of the topic “crude”. The raw data set was
preprocessed to first remove digits, punctuation marks, extra white spaces, then convert to lower
case, and remove stop words and stem words. It ends up with a \eqn{70\times 1517} document-term matrix,
where a row represents a document and a column represents a term and records the frequency of a
term. 

```{r}
data(acq)
data(crude)

coall <- c(acq,crude)
coall <- tm_map(coall, removeNumbers) # remove digits
coall <- tm_map(coall,removePunctuation)
coall <- tm_map(coall,content_transformer(tolower))
coall <- tm_map(coall,stripWhitespace)
coall <- tm_map(coall,removeWords,stopwords("english"))
coall <- tm_map(coall,removeWords,stopwords("SMART"))
coall <- tm_map(coall,removeWords,stopwords("catalan"))
coall <- tm_map(coall,removeWords,stopwords("romanian"))
coall <- tm_map(coall,removeWords,stopwords("german"))
coall <- tm_map(coall,stemDocument)
tdm <- DocumentTermMatrix(coall)
dim(tdm)

# Construct data set
tdm <- as.matrix(tdm)
constVar <- (apply(tdm,2,sd)==0)

#Remove constant variables
tdm <- tdm[,!constVar]

cl.tdm <- c(rep(1,50),rep(2,20))
n <- nrow(tdm)
p <- ncol(tdm)
```


## Evaluation
The performance of a classifier was assessed by the mean classification error rate estimated
by a 10-fold cross-validations with each fold containing 5 documents of the topic
“acq” and 2 documents of the topic “crude”.

The model tuning is done within each fold by an additional cross-validation.

```{r, warning=FALSE}
# Fix the CV folds
set.seed(193)
nfolds <- 10
foldid <- numeric(n)
K <- 2
for(k in 1:K){
  indexk <- cl.tdm==k
  nk <- sum(indexk)
  foldid[indexk] <- sample(rep(1:nfolds,ceiling(nk/nfolds))[1:nk],size = nk,replace = FALSE)
}

# Tuning setting (Use a sparse set for illustration purpose)
nfolds_tuning <- 4
ncpu <- 1
seed_tuning <- 123
thetaList <- matrix(rep(seq(0.3,0.9,0.1),p),ncol=p)
lambda <- c(1,0.1,0.01,0.001,0.0001)
cost <-  c(0.5,1,2)

cvErr <- data.frame(fold=1:nfolds, MC=0, QC=0, EQCLSVM=0, EQCRidge=0, EQCLasso=0)

for(i in 1:nfolds){
  cat(i,"")
  # Extract CV folds
  train <- tdm[foldid!=i,]
  test <- tdm[foldid==i,]
  cl.train <- cl.tdm[foldid!=i]
  cl.test <- cl.tdm[foldid==i]

  #---------Model Fitting--------------#
  # Median-based classifier (MC)
  MC <- eqcTrain(train,cl.train,
                 thetaList=matrix(rep(0.5,p),nrow = 1),
                 method = "qc")
  predMC <- predict(MC,newdata = test,type = "class")[[1]]
  cvErr$MC[i] <-  mean(predMC!=cl.test)

  # Quantile-based classifier (QC)
  QC <- eqcTrain(train,cl.train,
                 thetaList=thetaList,
                 method = "qc",
                 tuneControl = list(nfolds=nfolds_tuning, fold.seed=seed_tuning,ncpu=ncpu))
  predQC <- predict(QC,newdata = test,type = "class")[[1]]
  cvErr$QC[i] <-  mean(predQC!=cl.test)


  # EQC/LSVM
  EQCLSVM <- eqcTrain(train,cl.train,
                       thetaList=thetaList,
                       method = "svm",
                       kernel = "linear",cost = cost,
                       tuneControl = list(nfolds=nfolds_tuning, fold.seed=seed_tuning,ncpu=ncpu))
  predEQCLSVM <- predict(EQCLSVM,newdata = test,type = "class")[[1]]
  cvErr$EQCLSVM[i] <-  mean(predEQCLSVM!=cl.test)

  # EQC/Ridge
  EQCRidge <- eqcTrain(train,cl.train,
                       thetaList=thetaList,
                       method = "glmnet",
                       alpha = 0,lambda = lambda,
                       tuneControl = list(nfolds=nfolds_tuning, fold.seed=seed_tuning,ncpu=ncpu),
                       lower.limits=0, upper.limits=Inf)
  predEQCRidge <- predict(EQCRidge,newdata = test,type = "class")[[1]]
  cvErr$EQCRidge[i] <-  mean(predEQCRidge!=cl.test)


  # EQC/Lasso
  EQCLasso <- eqcTrain(train,cl.train,
                       thetaList=thetaList,
                       method = "glmnet",
                       alpha = 1,lambda = lambda,
                       tuneControl = list(nfolds=nfolds_tuning, fold.seed=seed_tuning,ncpu=ncpu),
                       lower.limits=0, upper.limits=Inf)
  predEQCLasso <- predict(EQCLasso,newdata = test,type = "class")[[1]]
  cvErr$EQCLasso[i] <-  mean(predEQCLasso!=cl.test)
}

cvErr

colMeans(cvErr[,-1])
```

## Reference
