---
title: "Minimal Simulation Study of the Ensemble Quantile Classifier"
author: "Yuanhao"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: EQC_July22.bib
vignette: >
  %\VignetteIndexEntry{Minimal Simulation Study of the Ensemble Quantile Classifier}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette introduces the functionality of the `eqc` package for 
fitting the ensemble quantile classifier (EQC).
As an example, a simulation study is done to compare the performance with regard to the test errors of the quantile-based classifier (QC) and the following methods.

* Ridge: Ridge logistic regression [@park2007penalized];
* EQC/Ridge: EQC fitted with the Ridge-type binomial loss;
* MC: Median-based classifier [@HALL2009].
* RSVM: SVM with radial kernel [@cortes1995support].

Tuning parameters were selected by minimizing the 5-fold cross validation errors, 
where the tuning set of $\theta$ is $\{0.02,0.04,\dotsc,0.98\}$,and
the tuning set of the penalty parameter of the ridge regression is $\{3,1,0.1,0.01,0.001,0.0001\}$.

```{r, warning=FALSE,message=FALSE}
library(glmnet)
library(e1071)
library(tidyr)
library(eqc)
```


## Heterogenous Skewed Data Generation
We assume an observation is from either Class 1 or Class 2.
We will investigate the heterogenous skewed scenario.
For each observation, 
it contains equal number of variables with the same distributions as
$W$, $\exp(W)$, $\log(|W|)$, $W^2$ and $|W|^{0.5}$, where $W ∼ N(0, 1)$.
In addition, each variable is further standarized to have a unit variance.
We add a shift $0.2$ to half of the observations to produce observations from Class 2 and
the remain will belong to Class 1.

For each simulation, 
we control the training sample size $ntrain$ and the test sample size $ntest$,
the number of covariates $p$ and the percentage of covariates that are Gaussian noises $percentNoise$.
A general function to simulate such data is defined below.
```{r genData}
#Transform variables from Gaussian random variables
g1 <- function(x){
  pInfo <- length(x)
  pw <- round(pInfo/5)
  c(x[1:pw],
    exp(x[(pw+1):(pw*2)])/2.161197,
    log(abs(x[(pw*2+1):(pw*3)]))/1.110405,
    x[(pw*3+1):(pw*4)]^2/1.413936,
    abs(x[(pw*4+1):pInfo])^0.5/0.3490925 )
}
g2 <- function(x){
  pInfo <- length(x)
  pw <- round(pInfo/5)
  c(x[1:pw],
    exp(x[(pw+1):(pw*2)])/2.161197,
    log(abs(x[(pw*2+1):(pw*3)]))/1.110405,
    x[(pw*3+1):(pw*4)]^2/1.413936,
    abs(x[(pw*4+1):pInfo])^0.5/0.3490925 ) + rep(0.2,pInfo)
}

genData <- function(ntrain=100,ntest=2000,
                    p=20,percentNoise=0,
                    g1,g2){
  pNoise <- ceiling(p*percentNoise)
  pInfo <- p-pNoise
  
  #----Training Set-----#
  xtrain <- matrix(rnorm(ntrain*p),ntrain,p)
  ntrainc1 <- round(ntrain*0.5)  
  xtrain[1:(ntrainc1),1:pInfo] <- t( apply(xtrain[1:(ntrainc1),1:pInfo],1,g1) )
  xtrain[(ntrainc1+1):ntrain,1:pInfo] <- t( apply(xtrain[(ntrainc1+1):ntrain,1:pInfo],1,g2))
  trainY <- rep(1:2, c(ntrainc1,ntrain-ntrainc1))
  colnames(xtrain) <- paste("x",1:p,sep="")
  
  #----Testing Set-----#
  xtest <- matrix(rnorm(ntest*p),ntest,p)
  ntestc1 <- round(ntest*0.5)
  xtest[1:(ntestc1),1:pInfo] <- t( apply(xtest[1:(ntestc1),1:pInfo],1,g1) )
  xtest[(ntestc1+1):ntest,1:pInfo] <- t( apply(xtest[(ntestc1+1):ntest,1:pInfo],1,g2))
  testY <- rep(1:2, c(ntestc1,ntest-ntestc1))
  colnames(xtest) <- paste("x",1:p,sep="")
  
  fulldata <- list(train=xtrain,
                   cl.train=trainY,
                   test=xtest,
                   cl.test=testY)
  return(fulldata)
}
```


## Simulation
We fix $ntrain=100$, $ntest=10000$ and $p=100$.
We let the number of Gaussian noise $percentNoise$ be within $\{0,0.5,0.9\}$ 
We repeat simulations of each combination of settings for $R=3$ times.

The following R script computes the test error rates for each method.
```{r simulation}

#Initial parameters
R <- 3  #Replicates
ntest<- 10000
ntrain<- 100
p <- 100
percentNoise <- c(0,0.5,0.9)

# Tuning setting
nfolds <- 4
ncpu <- 1
thetaList <- matrix(rep(seq(0.02,0.98,0.02),p),ncol=p)
lambda <- c(3,1,0.1,0.01,0.001,0.0001)
cost <- 2^(0:4)
gamma <- c(0.001, 0.01, 0.1, 1, 2)

# Fix cv folds for tuning parameters
set.seed(193)
n1 <- round(ntrain*0.5)
n2 <- ntrain-n1
index1 <- sample(rep(1:nfolds,ceiling(n1/nfolds))[1:n1],size = n1,replace = FALSE)
index2 <- sample(rep(1:nfolds,ceiling(n2/nfolds))[1:n2],size = n2,replace = FALSE)
foldid <- c(index1,index2)
testErr <- data.frame(QC=0,Ridge=0, EQCRidge=0, MC=0, RSVM=0,
                      percentNoise=rep(c(0,0.5,0.9),each=R))

cat("noiselevel:")
for(noiselevel in 1:3){
  cat(noiselevel,"")
  for(i in 1:R){
    # Generate data
    fulldata <- genData(ntrain,ntest,
                        p=p,percentNoise=percentNoise[noiselevel],
                        g1,g2)
    
    train <- fulldata$train
    test <- fulldata$test
    cl.train <- fulldata$cl.train
    cl.test <- fulldata$cl.test
    
    #---------Model Fitting--------------#
    # Ridge
    Ridge <- cv.glmnet(x = train,y = as.factor(cl.train),
                       lambda=lambda,alpha=0,
                       nfolds = nfolds,foldid=foldid,
                       family = "binomial")
    predRidge <- ifelse( predict(Ridge,newx = test,type = "response", s="lambda.1se")<0.5,1,2)
    testErr$Ridge[R*(noiselevel-1)+i] <- mean(predRidge!=cl.test)
    
    # EQC/Ridge
    EQCRidge <- eqcTrain(train,cl.train,
                         thetaList=thetaList,
                         method = "glmnet",
                         alpha = 0,lambda = lambda,
                         tuneControl = list(foldid=foldid,ncpu=ncpu),
                         lower.limits=0, upper.limits=Inf)
    predEQCRidge <- predict(EQCRidge,newdata = test,type = "class")[[1]]
    testErr$EQCRidge[R*(noiselevel-1)+i] <-  mean(predEQCRidge!=cl.test)
    
    # Quantile-based classifier (QC)
    QC <- eqcTrain(train,cl.train,
                   thetaList=thetaList,
                   method = "qc",
                   tuneControl = list(foldid=foldid))
    predQC <- predict(QC,newdata = test,type = "class")[[1]]
    testErr$QC[R*(noiselevel-1)+i] <-  mean(predQC!=cl.test)
    
    
    # Median-based classifier (MC)
    MC <- eqcTrain(train,cl.train,
                   thetaList=matrix(rep(0.5,p),nrow = 1),
                   method = "qc")
    predMC <- predict(MC,newdata = test,type = "class")[[1]]
    testErr$MC[R*(noiselevel-1)+i] <-  mean(predMC!=cl.test)
    
    # RSVM
    rsvmfitTune <- tune.svm(x = train,y = factor(cl.train,levels = c(1,2)),
                            type="C-classification", gamma = gamma, cost = cost)
    rsvmfit <- svm(x = train,y = factor(cl.train,levels = c(1,2)),
                   type="C-classification",
                   gamma=rsvmfitTune$best.parameters[1],cost=rsvmfitTune$best.parameters[2])
    predrsvm <- predict(rsvmfit,newdata = test)
    testErr$RSVM[R*(noiselevel-1)+i] <- mean(predrsvm!=factor(cl.test,levels = c(1,2)))
  }
}

```

We then reformat the test error of each method.
```{r relErr}
testErr_reformat <- tidyr::gather(testErr,key = "method",value = "testError",-percentNoise)
testErr_reformat$method <- ordered(testErr_reformat$method, levels=c("QC","Ridge","EQCRidge","MC","RSVM"))
testErr_reformat$percentNoise <- factor(testErr_reformat$percentNoise)
library(latex2exp) 
levels(testErr_reformat$percentNoise) <-  TeX( c("$Noise = 0%$","$Noise = 50%$","$Noise = 90%$") )
head(testErr_reformat)
```

## Visualization
```{r, fig.cap = "Boxplots of Test Errors of each Method", fig.width=7,fig.height=3}
library(ggplot2)
ggplot(testErr_reformat, mapping=aes(x=method, y=testError)) +
  geom_boxplot(outlier.alpha = 0.5) +
  geom_hline(yintercept = 0,linetype="dashed") +
  labs(x="Method",y="Test error")+
  theme(legend.position="top",text = element_text(size=15) )+
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  facet_grid(~percentNoise, labeller = label_parsed)
```


## Reference
